---
title: freddie mac
output: html_document
---
##data explore
```{r}
load("~/Desktop/Freddie Mac data/USMortgages2008_2009.rdata")
names(D1)
```

use 10% data as traning data here

```{r}
sample_index = sample(1:nrow(D1), floor(nrow(D1)/10), replace=FALSE)
D2 <- D1[sample_index,]
D3 <- D1[-sample_index,]

head(D2)
head(D1)

summary(D1$def_flag)
summary(D2$def_flag)
```

extract the numeric variables

```{r}
D2_num = subset(D2, select=c("score","CLTV","DTI",
                             "UPB","LTV","OIR","def_flag"))

D3_num = subset(D3, select=c("score","CLTV","DTI",
                             "UPB","LTV","OIR","def_flag"))
library(car)
scatterplot.matrix(~score+CLTV+DTI+UPB+LTV+OIR, data=D2_num[1:1000,],
                   main="Scatterplot Matrix",pch=".")

D2_def = D2[D2$def_flag == TRUE,]
scatterplot.matrix(~score+CLTV+DTI+UPB+LTV+OIR, data=D2_def,
                   main="Scatterplot Matrix_default data",pch=".")
```

LTV CLTV show collinearity
different densoty plot between two classes can be seen on score DTI OIR

##logistic regression
#remove NA
Simply remove the rows with "NA" in it. Maybe it is better to use some strategy to  handle with missing data problem.

```{r}
temp = D2_num[complete.cases(D2_num),]
summary(temp)
```

#build a test set
```{r}
test = D3_num[complete.cases(D3_num),]
```

data prepared

#using function glm
fit a logistic model with two variables

```{r}
library(Deducer)

glm.fit = glm(def_flag ~ DTI + OIR, data=temp, family=binomial)
summary(glm.fit)
coef(glm.fit)
glm.probs=predict(glm.fit,type="response")
sum(glm.probs>0.5);sum(glm.probs>0.1)

library(ROCR)
pr <- prediction(glm.probs, temp$def_flag)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]];auc
plot(prf);lines(x = c(0,1), y = c(0,1))
```

#use all numeric variable
```{r}
glm.fit2 = glm(def_flag ~ score+CLTV+DTI+UPB+LTV+OIR, data=temp, family=binomial)
summary(glm.fit2)
coef(glm.fit2)
glm.probs=predict(glm.fit2,type="response")
sum(glm.probs>0.5);sum(glm.probs>0.1)

pr <- prediction(glm.probs, temp$def_flag)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]];auc
plot(prf);lines(x = c(0,1), y = c(0,1))
```

#on test dataset
```{r}
fitted.results <- predict(glm.fit2,newdata=test,
                          type='response')
fitted.outputs <- ifelse(fitted.results > 0.5, 1, 0)
misClasificError <- mean(fitted.outputs != test$def_flag,na.omit="TRUE")
misClasificError

threshold = function(x){
  fitted.outputs <- ifelse(fitted.results > x, 1, 0)
  misClasificError <- mean(fitted.outputs != test$def_flag,na.omit="TRUE")
  return (misClasificError)
 }

rate = lapply(seq(0.1,0.9,0.1),threshold)
unlist(rate)

fitted.outputs <- ifelse(fitted.results > 0.5, 1, 0)
table(fitted.outputs,test$def_flag)

fitted.outputs <- ifelse(fitted.results > 0.01, 1, 0)
table(fitted.outputs,test$def_flag)
```

question about how to choose threshold?

```{r}
as.numeric(summary(D1$def_flag)[3])/nrow(D1)

pr <- prediction(fitted.results, test$def_flag)
prf <- performance(pr, measure = "tpr", x.measure = "fpr")
plot(prf);lines(x = c(0,1), y = c(0,1))

auc <- performance(pr, measure = "auc")
auc <- auc@y.values[[1]]
auc

```

##try LDA
```{r}
library (MASS)
lda.fit=lda(def_flag ~ score+CLTV+DTI+UPB+LTV+OIR, data=temp)
lda.fit
lda.pred = predict (lda.fit , newdata=test, type='response')
names(lda.pred)
lda.class =lda.pred$class
table(lda.class,test$def_flag )
```

##try to plus other variables into logistic model
#maybe logistic model plus lasso will be useful, glmnet and glmpath
```{r}
load("~/Desktop/Freddie Mac data/USMortgages2008_2009.rdata")
names(D1)
head(D1)
#first.time.homebuyer,insurance,number.units,occupance.status,channle,PPM
#product.type,property.state,property.type,loan.purpose,orig.loan.term
#number.borrowers,seller,servicer,loan_age 
factor_data = subset(D1, select=c("first.time.homebuyer","insurance","number.units",
                             "occupancy.status","channel","PPM","product.type",
                             "property.type","loan.purpose","number.borrowers",
                             "seller","servicer","def_flag"))

element = sapply(factor_data,unique);element
count_factor = sapply(factor_data,function(x) {table(x,exclude = NULL)});count_factor
```

